Landscape composition and beneficial insects - time is important

### 2020-08-05 manuscript/manuscript_outline.Rmd
Removed flax from models. Updated figures.

@@

To characterize landscape composition surrounding the traps, we used classified landscape data from @AAFC_data.
<!-- + AAFC cropland landscape classifications are very accurate for cultivated fields (~90%), but are less accurate for non-crop areas (~70%) -->
We combined functionally similar landscape categories (cereals: _Triticum aestivum_ and _Hordeum vulgare_, pulses: _Pisum sativum_ and _Lens culinaris_, forest: coniferous and broadleaf), then extracted the landscape composition within a circle of a fixed radius surrounding each trap location. 
However, the radius of the circle necessary to characterize the landscape depends on the ``grain" of the landscape that is relevant to the abundance of a given organism [@addicott1987].
Therefore, we divided the landscape into annuli (rings) surrounding each trap location.
Each annulus was 30 m thick, and had inner radii ranging from 30 to 1470m, in increments of 30m.
Using functional regression [@ramsay2004], we were able to incorporate landscape information at different distance, allowing us to assess the impact of both local and regional landscape composition.

@@

Activity density of each arthropod was fit as generalized additive models @wood2017, using _mgcv_ version `r packageVersion('mgcv')`. 
Count data of arthropods were fit using a negative binomial distribution with a log-link.
To account for different lengths of trap exposure, log-days since trap placement was used as a fixed effect with its slope held at 1 ("offset" variable).
We included day of year as a 1-dimensional smooth and easting and northing as a 2-dimensional smooth, in order to account for underlying spatial and temporal autocorrelation in the data.
For each landscape term, we used three SoF regression terms:
First, we used the proportion cover of all landscape cover classes in the annuli surrounding each trap location (spatial effect of cover class).
Second, we used day of the year multiplied by the proportion of cover classes in each annuli (temporal effect of cover class). 
Finally, we used a tensor-product interaction of landscape cover and day of year (spatio-temporal interaction of cover class).
We used 10 basis dimensions for the spatial and temporal landscape smoothers, and 16 basis dimensions for the spatio-temporal interaction.
This was deemed adequate, as the effective degrees of freedom for all smoothing terms were far below the number of basis dimensions, indicating that no additional basis dimensions were needed (see Tables \@ref(tab:PteMelSmooth), \@ref(tab:ParDisSmooth), \@ref(tab:ParMoeSmooth), \@ref(tab:OpilioSmooth)).
To remove unimportant terms from the model, we used thin-plate regression splines with shrinkage ("bs=`ts'" in _mgcv_).
This is a continuous analog to stepwise model selection, whereby weaker terms can be completely removed from the model, rather than just reduced to a line [@marra2011].
<!-- + Double-penalization gave qualitatively similar results, but required twice the number of penalty terms, and some of the penalty terms were unstable  -->

@@

Some of the landscape terms in our model were strongly concurved, so we removed them from the model or combined them.
Concurvity is a nonlinear analogue of multicollinearity, and can bias estimates of standard errors [@buja1989].
There is no agreed-upon threshold of "unacceptable" concurvity, but 0.5 is commonly used [@dominici2002; @ramsay2003]. 
Some landscape terms were strongly concurved (Figure \@ref(fig:concurvity)), namely forest and shrubland (max: 0.71), canola and cereal (max: 0.74), and grassland and wetland (max: 0.63).
Forest and shrubland were combined into the ``Tree/Shrub" category, grassland and wetland were combined into the "Grass/Wetland" category, as they were strongly positively correlated at all distances, indicating similar feature classes.
Cereal was removed from the model, as it was negatively correlated with canola cover at distances less than 200m and was positively correlated at distances over 500m, reflecting the most common crop rotations (canola -> wheat -> barley) and the landscape structure present across our study region.

@@

+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Cover class     | _P. melanarius_            | _P. distincta_             | _P. moesta_                 | _P. opilio_           |
+=================+============================+============================+=============================+=======================+
| Grass & Wetland | - Local sink (late)        |                            | - Local sink                | - Local sink (early)  |
|                 |                            |                            | - Landscape source          |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Canola          | - Local source (early)     | - _Local source (early)_   | - Local source (early)      |                       |
|                 | - Landscape sink (early)   | - _Local sink (late)_      | - Local sink (late)         |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Pasture         |                            | - Landscape sink           |                             |                       | 
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Forest & Shrubs |                            | - Overall sink (late)      | - _Overall source (early)_  | - Local sink (early)  |
|                 |                            |                            | - _Overall sink (late)_     |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Pulses          | - Overall source (late)    |                            | - Landscape source          |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Flax            | - Local sink               |                            |                             | - Overall sink        |
|                 | - Landscape sink           |                            |                             |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Urban           |                            |                            | - Local source (early)      |                       |
|                 |                            |                            | - Local sink (early)        |                       |
|                 |                            |                            | - Landscape sink (late)     |                       |
|                 |                            |                            | - Landscape source (late)   |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+

@@

```{r concurvity, echo=FALSE, fig.cap='Concurvity estimates for top land cover classes'}  
  include_graphics('../figures/coverCorPlots/concurvityEstimate.png')
```

```{r landscapeComp, echo=FALSE, fig.cap='Landscape composition in annuli surrounding each trap location'}  
  include_graphics('../figures/oRingPlot2_AAFC.png')
```

>>>>

#Get landscape % cover for each cover class
coverList <- PteMelMod$datList[9:25]
siteID <- as.numeric(rownames(unique(cbind(PteMelMod$datList$E,PteMelMod$datList$N))))
coverList <- lapply(coverList,function(x){
  data.frame(site=as.character(siteID),x[siteID,]) %>%  #Get unique sites only, convert to DF
  pivot_longer(cols=d30:d1500,names_to='dist',values_to='prop') %>% 
  mutate(dist=as.numeric(gsub('d','',dist)))
})
coverList <- data.frame(cover=rep(names(coverList),each=nrow(coverList[[1]])),do.call('rbind',coverList))

@@

To characterize landscape composition surrounding the traps, we used classified landscape data from @AAFC_data.
<!-- + AAFC cropland landscape classifications are very accurate for cultivated fields (~90%), but are less accurate for non-crop areas (~70%) -->
We combined functionally similar landscape categories (cereals: _Triticum aestivum_ and _Hordeum vulgare_, pulses: _Pisum sativum_ and _Lens culinaris_, forest: coniferous and broadleaf), then extracted the landscape composition within a circle of a fixed radius surrounding each trap location. 
The ten most-common cover classes represented 98% of the total land cover in our study region: grassland, cereal, canola, pasture, pulses, wetland, urban, shrubland, flax, and forest (Figure \@ref(fig:landscapeComp)); flax was removed, as only a single site had nearby flax cover.
The radius of the circle necessary to characterize the landscape depends on the scale of the landscape that matters to a given organism [@addicott1987].
Therefore, we divided the landscape into annuli (rings) surrounding each trap location.
Each annulus was 30 m thick, and had inner radii ranging from 30 to 1470m, in increments of 30m.
Using functional regression [@ramsay2004], we were able to incorporate landscape information at different distance, allowing us to assess the impact of both local and regional landscape composition.

@@

Activity density of each arthropod was fit as generalized additive models @wood2017, using _mgcv_ version `r packageVersion('mgcv')`. 
Count data of arthropods were fit using a negative binomial distribution with a log-link.
To account for different lengths of trap exposure, log-days since trap placement was used as a fixed effect with its slope held at 1 ("offset" variable).
We included day of year as a 1-dimensional smooth and easting and northing as a 2-dimensional smooth, in order to account for underlying spatial and temporal autocorrelation in the data.
The deviance residuals from each model were visually inspected for normality and equal variance.

For each landscape term, we used three SoF regression terms:
First, we used the proportion cover of all landscape cover classes in the annuli surrounding each trap location (spatial effect of cover class).
Second, we used day of the year multiplied by the proportion of cover classes in each annuli (temporal effect of cover class). 
Finally, we used a tensor-product interaction of landscape cover and day of year (spatio-temporal interaction of cover class).
We used 10 basis dimensions for the spatial and temporal landscape smoothers, and 16 basis dimensions for the spatio-temporal interaction.
This was deemed adequate, as the effective degrees of freedom for all smoothing terms were far below the number of basis dimensions, indicating that no additional basis dimensions were needed (see Tables \@ref(tab:PteMelSmooth), \@ref(tab:ParDisSmooth), \@ref(tab:ParMoeSmooth), \@ref(tab:OpilioSmooth)).
To remove unimportant terms from the model, we used thin-plate regression splines with shrinkage ("bs=`ts'" in _mgcv_).
This is a continuous analog to stepwise model selection, whereby weaker terms can be completely removed from the model rather than reduced to a line [@marra2011].
<!-- + Double-penalization gave qualitatively similar results, but required twice the number of penalty terms, and some of the penalty terms were unstable  -->

@@

Some of the landscape terms in our model were strongly concurved, so we removed them from the model or combined them.
Concurvity is a nonlinear analogue of multicollinearity, and can bias estimates of standard errors [@buja1989].
There is no agreed-upon threshold of "unacceptable" concurvity, but 0.5 is commonly used [@dominici2002; @ramsay2003]. 
Some landscape terms were strongly concurved (Figure \@ref(fig:concurvityFull)), namely forest and shrubland (max: 0.71), canola and cereal (max: 0.74), and grassland and wetland (max: 0.63).
Forest and shrubland were combined into the ``Tree/Shrub" category, grassland and wetland were combined into the "Grass/Wetland" category, as they were strongly positively correlated at all distances, indicating similar feature classes.
Cereal was removed from the model, as it was negatively correlated with canola cover at distances less than 200m and was positively correlated at distances over 500m, reflecting the most common crop rotations (canola -> wheat -> barley) and the landscape structure present across our study region.
This resulted in a much-reduced set of landscape terms, none of which were strongly concurved with each other (Figure \@ref(fig:concurvity))

@@

+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Cover class     | _P. melanarius_            | _P. distincta_             | _P. moesta_                 | _P. opilio_           |
+=================+============================+============================+=============================+=======================+
| Grass & Wetland | - Local sink (late)        |                            | - Local sink                | - Local sink (early)  |
|                 |                            |                            | - Landscape source          |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Canola          | - Local source (early)     | - _Local source (early)_   | - Local source (early)      |                       |
|                 | - Landscape sink (early)   | - _Local sink (late)_      | - Local sink (late)         |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Pasture         |                            | - Landscape sink           |                             |                       | 
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Forest & Shrubs |                            | - Overall sink (late)      | - _Overall source (early)_  | - Local sink (early)  |
|                 |                            |                            | - _Overall sink (late)_     |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Pulses          | - Overall source (late)    |                            | - Landscape source          |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+
| Urban           |                            |                            | - Local source (early)      |                       |
|                 |                            |                            | - Local sink (early)        |                       |
|                 |                            |                            | - Landscape sink (late)     |                       |
|                 |                            |                            | - Landscape source (late)   |                       |
+-----------------+----------------------------+----------------------------+-----------------------------+-----------------------+

@@

```{r landscapeComp, echo=FALSE, fig.cap='Landscape composition in annuli surrounding each trap location. Sites are represented by individual lines. Overall percent cover for each cover class is listed in each sub-figure heading.'}  
coverList %>% group_by(cover) %>% mutate(p=sum(prop)) %>% ungroup() %>% mutate(p=100*round(p/sum(unique(p)),4)) %>% 
  mutate(cover=paste0(as.character(cover),' (',p,'%)'),prop=prop*100) %>% 
  mutate(cover=factor(cover,levels=unique(cover))) %>% 
  ggplot(aes(dist,prop,group=site))+geom_line(alpha=0.2)+
  facet_wrap(~cover)+theme_classic()+labs(y='Percent cover',x='Distance (m)')

```{r concurvityFull, echo=FALSE, fig.cap='Concurvity estimates for ten most-abundant land cover classes'}  
  include_graphics('../figures/coverCorPlots/concurvityEstimate_full.png')
```

```{r concurvity, echo=FALSE, fig.cap='Concurvity estimates for reduced cover classes used in models'}  
  include_graphics('../figures/coverCorPlots/concurvityEstimate_reduced.png')
```
